---
title: "Introduction"
author: "Jonathan M. Hill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[![codecov](https://codecov.io/github/ficonsulting/RInno/branch/master/graphs/badge.svg)](https://codecov.io/github/ficonsulting/RInno)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

RInno makes it easy to install local shiny apps by providing an interface between R and [Inno Setup](http://www.jrsoftware.org/isinfo.php), an installer for Windows programs (sorry Mac and Linux users). It is designed to be simple to use (two lines of code at a minimum), yet comprehensive.

If a user does not have R installed, the RInno installer can be configured to ask them to install R along with a shiny app, `include_R = TRUE`. And similar to Dr. Lee Pang's [DesktopDeployR](https://github.com/wleepang/DesktopDeployR) project, RInno provides a framework for managing software dependencies and error logging features. However, RInno also supports GitHub package dependencies, continuous installation (auto-update on start up), and it is easier to manage with `create_app`, the main RInno function. DesktopDeployR requires many manual adjustments and a deep understanding of the entire framework to use, but RInno can be learned incrementally and changes automatically flow down stream. You don't need to remember the 100+ places impacted by changing `app_dir`. It only requires a high-level understanding of what you'd like to accomplish, and RInno takes care of the rest.

## Getting Started

```{r, eval=FALSE}
# Require Package
require(RInno)

# Use RInno to get Inno Setup
RInno::install_inno()
```

If you don't use [installr](https://github.com/talgalili/installr) to update your R version, you should. The next release of installr will include `install.inno` along with a lot of other cool stuff. After that release, `install_inno` will no longer be supported here.

## Minimal example

Once you have developed a shiny app, you can build an installer with `create_app` followed by `compile_iss`.

```{r, eval=FALSE}
# Example app included with RInno package
example_app(wd = getwd())

# Build an installer
create_app(app_name = "Your appname", app_dir = "app")
compile_iss()
```

`create_app` creates an installation framework in your app's directory, `app_dir`.  Feel free to customize it before you call `compile_iss`.  For example, you can replace the default/setup icon at [Flaticon.com](http://www.flaticon.com/), or you can customize the pre-install message, *infobefore.txt*. Just remember, the default values for those files have not changed. The Inno Setup Script (ISS), *app_name.iss*, will look for *default.ico* and use it as your app's icon. You must replace *default.ico*, or rename *default.ico* in the ISS to use a different one.

For custom installations, see below.

## server.R Requirements

In order to close the app when your user's session completes:

1. Add `session` to your `server` function
2. Call `stopApp()` when the session ends 

```{r, eval=FALSE}
function(input, output, session) {

  session$onSessionEnded(function() {
      stopApp()
      q("no")
  })
}
```


If you forget to do this, users will complain that their icons are broken and rightly blame you for it (an R session will be running in the background hosting the app, but they will need to press ctrl + alt + delete and use their task manager to close it). **Not cool**.

## Custom Installer

If you would like to create a custom installer from within R, you can slowly build up to it with `create_app`, like this:

```{r, eval=FALSE}
create_app(
  app_name    = "My AppName", 
  app_dir     = "My/app/path",
  dir_out     = "wizard",
  pkgs        = c("jsonlite", "shiny", "magrittr", "xkcd"),  # CRAN-like repo packages
  remotes     = c("talgalili/installr", "daattali/shinyjs"), # GitHub packages
  include_R   = TRUE,   # Download R and install it with your app, if necessary
  R_version   = 2.2.1,  # Old versions of R
  privilege   = "high", # Admin only installation
  default_dir = "pf")   # Install app in to Program Files
```

`create_app` passes its arguments to most of the other support functions in RInno, so you can specify most things there and they will get passed on, or you can provide detailed instructions directly to those functions like this:

```{r, eval=FALSE}
# Copy installation scripts (JavaScript, icons, readme.txt, run.R, app.R)
copy_installation(app_dir = "my/app/path")

# If your users need R installed:
get_R_exe(app_dir = "my/app/path", R_version = 2.2.1)

# Create batch file
create_bat(app_name = "My AppName", app_dir = "my/app/path")

# Create app config file
create_config(app_name = "My AppName", R_version = 2.2.1, app_dir = "my/app/path",
  pkgs = c("jsonlite", "shiny", "magrittr", "dplyr", "caret", "xkcd"))

# Build the iss script
start_iss(app_name = "My AppName") %>%

  # C-like directives
  directives(R_version   = 2.2.1, 
             include_R   = TRUE, 
             app_version = "0.1.2",
             publisher   = "Your Company", 
             main_url    = "yourcompany.com") %>%

  # Setup Section
  setup(output_dir  = "wizard", 
        app_version = "0.1.2",
        default_dir = "pf", 
        privilege   = "high",
        inst_readme = "pre-install instructions.txt", 
        setup_icon  = "myicon.ico",
        pub_url     = "mycompany.com", 
        sup_url     = "mycompany.github.com/issues",
        upd_url     = "mycompany.github.com") %>%

  # Languages Section
  languages() %>%

  # Tasks Section
  tasks(desktop_icon = FALSE) %>%

  # Files Section
  files(app_dir = "my/app/path", file_list = "path/to/extra/files") %>%

  # Icons Section
  icons(app_desc       = "This is my local shiny app",
        app_icon       = "notdefault.ico",
        prog_menu_icon = FALSE,
        desktop_icon   = FALSE) %>%

  # Execution & Pascal code to check registry during installation
  # If the user has R, don't give them an extra copy
  # If the user needs R, give it to them
  run() %>%
  code() %>%

  # Write the Inno Setup script
  writeLines(file.path("my/app/path", "My AppName.iss"))

  # Check your files, then
  compile_iss()

```

Feel free to read the Inno Setup [documentation](http://www.jrsoftware.org/ishelp/) and RInno's documentation to get a sense for what is possible. Also, please suggest useful features or build them yourself! We have a very positive culture at FI Consulting, and we would love to get your feedback.

Please note that this project has a [Contributor Code of Conduct](https://github.com/ficonsulting/RInno/blob/master/CONDUCT.md). By participating in this project you agree to abide by its terms.

## License

The RInno package is licensed under the GPLv3. See LICENSE for details.
