---
title: "Introduction"
author: "Jonathan M. Hill"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[![codecov](https://codecov.io/github/ficonsulting/RInno/branch/master/graphs/badge.svg)](https://codecov.io/github/ficonsulting/RInno)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

RInno makes it easy to install local shiny apps by providing an interface between R and [Inno Setup](http://www.jrsoftware.org/isinfo.php), an installer for Windows programs (sorry Mac and Linux users). It is designed to be simple to use (two lines of code at a minimum), yet comprehensive.

If a user does not have R installed, the RInno installer can be configured to ask them to install R along with a shiny app, `include_R = TRUE`. And similar to Dr. Lee Pang's [DesktopDeployR](https://github.com/wleepang/DesktopDeployR) project, RInno provides a framework for managing software dependencies and error logging features. However, RInno also supports GitHub package dependencies, continuous installation (auto-update on start up), and it is easier to manage with `create_app`, the main RInno function. DesktopDeployR requires many manual adjustments and a deep understanding of the entire framework to use, but RInno can be learned incrementally and changes automatically flow down stream. You don't need to remember the 100+ places impacted by changing `app_dir`. RInno only requires a high-level understanding of what you'd like to accomplish.

## Getting Started

```{r, eval=FALSE}
# Require Package
require(RInno)

# Use RInno to get Inno Setup
RInno::install_inno()
```

## Minimal example

Once you have developed a shiny app, you can build an installer with `create_app` followed by `compile_iss`.

```{r, eval=FALSE}
# Example app included with RInno package
example_app(wd = getwd())

# Build an installer
create_app(app_name = "Your appname", app_dir = "app")
compile_iss()
```

`create_app` creates an installation framework in your app's directory, `app_dir`.  You can perform minor customizations before you call `compile_iss`.  For example, you can replace the default/setup icon at [Flaticon.com](http://www.flaticon.com/), or you can customize the pre-/post- install messages, *infobefore.txt* and *infoafter.txt*. Just remember, the default values (i.e. `create_app(info_after = "infobefore.txt")`) for those files have not changed. The Inno Setup Script (ISS), *app_name.iss*, will look for *default.ico* and try to use it until you update the script or call `create_app` with the new icon's file name (i.e. `create_app(app_icon = "new.ico")`).

Chrome is the default browser used by RInno because of its app mode feature and development-minded features. IE/Edge often prevents icons and third party JavaScript libraries from loading because of various IT policies, which can result in bugs. The default `user_browser` setting will open Chrome in app mode, which looks more like a stand-alone app than when it opens in another tab of your default browser. Regardless of which browser you specify, RInno's startup sequence will fall back on the user's default browser if it is not installed.

## ui.R Requirements

In order to replace Chrome's logo with your app's icon, add something like this to your ui.R file:

```
fluidPage(
  tags$head(
    tags$link(
      rel = "icon", 
      type = "image/x-icon", 
      href = "http://localhost:1984/default.ico")
  )
)
```

## server.R Requirements

In order to close the app when your user's session completes:

1. Add `session` to your `server` function
2. Call `stopApp()` when the session ends 

```{r, eval=FALSE}
function(input, output, session) {

  session$onSessionEnded(function() {
      stopApp()
      q("no")
  })
}
```

If you forget to do this, users will complain that their icons are broken and rightly blame you for it (an R session will be running in the background hosting the app, but they will need to press ctrl + alt + delete and use their task manager to close it). **Not cool**.

## Package Dependency Management

Provide a list of packages to `create_app`, and RInno will install them with your shiny app. If the list does not have a version #, the default is `utils::packageVersion` to determine which version is installed in the development environment. Whereas if a specific package version is specified, i.e. `pkgs = c(shiny = "==1.0.5", "jsonlite", "httr")`, RInno will respect that specification and use `utils::packageVersion` for the rest.

```
create_app(
  app_name = "myapp", 
  app_dir = "app",
  pkgs = c("shiny", "jsonlite", "httr")
)
```
RInno will default to `shiny = paste0(">=", utils::packageVersion("shiny"))`, etc. if versions are not included. This provides a similar management strategy as developers use in the DESCRIPTION file's `Imports:` section of an R package. For more information, see [R Packages - Package Metadata](http://r-pkgs.had.co.nz/description.html) by Hadley Wickham.

## Custom Installations

If you would like to create a custom installer from within R, you can slowly build up to it with `create_app`, like this:

```{r, eval=FALSE}
create_app(
  app_name    = "My AppName", 
  app_dir     = "My/app/path",
  dir_out     = "wizard",
  pkgs        = c("jsonlite", "shiny", "magrittr", "xkcd"),  # CRAN-like repo packages
  remotes     = c("talgalili/installr", "daattali/shinyjs"), # GitHub packages
  include_R   = TRUE,     # Download R and install it with your app, if necessary
  R_version   = "2.2.1",  # Old versions of R
  privilege   = "high",   # Admin only installation
  default_dir = "pf")     # Install app in to Program Files
```

`create_app` passes its arguments to most of the other support functions in RInno, so you can specify most things there and they will get passed on, or you can provide detailed instructions directly to those functions like this:

```{r, eval=FALSE}
# Copy installation scripts (JavaScript, icons, infobefore.txt, package_manager.R, app.R)
copy_installation(app_dir = "my/app/path")

# If your users need R installed:
get_R(app_dir = "my/app/path", R_version = 2.2.1)

# Create batch file
create_bat(app_name = "My AppName", app_dir = "my/app/path")

# Create app config file
create_config(app_name = "My AppName", R_version = "2.2.1", app_dir = "my/app/path",
  pkgs = c("jsonlite", "shiny", "magrittr", "dplyr", "caret", "xkcd"))

# Build the iss script
start_iss(app_name = "My AppName") %>%

  # C-like directives
  directives_section(R_version   = "2.2.1", 
             include_R   = TRUE, 
             app_version = "0.1.2",
             publisher   = "Your Company", 
             main_url    = "yourcompany.com") %>%

  # Setup Section
  setup_section(output_dir  = "wizard", 
        app_version = "0.1.2",
        default_dir = "pf", 
        privilege   = "high",
        inst_readme = "pre-install instructions.txt", 
        setup_icon  = "myicon.ico",
        pub_url     = "mycompany.com", 
        sup_url     = "mycompany.github.com/issues",
        upd_url     = "mycompany.github.com") %>%

  # Languages Section
  languages_section() %>%

  # Tasks Section
  tasks_section(desktop_icon = FALSE) %>%

  # Files Section
  files_section(app_dir = "my/app/path", file_list = "path/to/extra/files") %>%

  # Icons Section
  icons_section(app_desc       = "This is my local shiny app",
        app_icon       = "notdefault.ico",
        prog_menu_icon = FALSE,
        desktop_icon   = FALSE) %>%

  # Execution & Pascal code to check registry during installation
  # If the user has R, don't give them an extra copy
  # If the user needs R, give it to them
  run_section() %>%
  code_section() %>%

  # Write the Inno Setup script
  writeLines(file.path("my/app/path", "My AppName.iss"))

  # Check your files, then
  compile_iss()

```

Feel free to read the Inno Setup [documentation](http://www.jrsoftware.org/ishelp/) and RInno's documentation to get a sense for what is possible. Also, please suggest useful features or build them yourself! We have a very positive culture at FI Consulting, and we would love to get your feedback.

Please note that this project has a [Contributor Code of Conduct](https://github.com/ficonsulting/RInno/blob/master/CONDUCT.md). By participating in this project you agree to abide by its terms.

## License

The RInno package is licensed under the GPLv3. See LICENSE for details.
